The abundance of Bullet-Groups in $\Lambda$CDM

### 2014-03-11 paper/ms.tex
section 3

Since then, other examples of Bullet-like systems have been found; 
MACS J0025.4-1222 \citep{Bradac2008}, Abell 2744 \citep{Merten2011},
DLSCL J0916.2+2951 \citep{Dawson2012}, ZwCl 1234.0+02916
\citep{Dahle2013}. Recently \citep{Gastaldello} observed baryonic-DM
displacement if $124\pm 20$ kpc in a group-like system with a total
mass $2.4\pm 0.6 \times 10^{14}$\Msun. Systems of this mass are
$\sim10$ times more abundant than cluster systems in the mass range of
the Bullet Cluster $>10^{15}$\hMsun, this opens up the possibility of
observationally finding bullet groups in a fair amount to impose
constraints on $\Lambda$CDM.

@@

The Bullet groups is composed by two halos. the host halo and the sub
halo. This configuration is described by the position and velocity of
the sub-halo in a frame of reference where the main halo is at rest;
thus  $\vec{v{}}=\vec{v}_{sub}-\vec{v}_{halo}$ and 
$\vec{r}=\vec{r}_{sub}-\vec{r}_{halo}$, where the subscripts $host$
and $sub$ refer to the host and sub-halo in the frame of reference of
the simulation, respectively.

The angle between these two vectors characterized by, 
\begin{equation}
  \mu\equiv
  \cos(\theta)=\frac{\vec{v{}}\cdotp{}\vec{r}}{\left\|\vec{v}{}\right\|
    \left\|\vec{r}\right\|} 
 \end{equation} 
%
encodes the geometry of the collision, i.e. cases of $|\mu|\approx 1$
can be considered as head-on collisions while $|\mu|\approx 0$
describe a grazing trajectory.

The bullet-like encounter can be instantaneously described by
quantities the following quantities the circular velocity of the host
and the sub-halo, $V_{\rm c,host}$ and $V_{\rm c,  sub}$; the size of
the host halo $R_{\rm vir}$; the relative position and velocity of the
substructure, $\vec{v}$ and $\vec{r}$; and the angle between the
position and velocity $\mu$.

As a first approximation there are two quantities that are available
from observations of these Bullet-like systems. The projected distance
between two dominant dark matter clumps and the ratio of the
galaxies luminosities associated to them. From the simulation point
of view this can be translated into the 2D projected values of
$||\vec{r}||$, its value relative to the virial radius $D_{\rm off}=
||\vec{r}||_{\rm 2D}/R_{\rm vir}$ and the ratio of the circular velocities
of the two clumps $V_{\rm c,sub}/V_{\rm c,host}$. In a higher degree
of detail, in order to gain better insight we use
the sub-structure velocity as a fraction of the host's circular
velocity, $||\vec{r}||/V_{\rm c, host}$, as a measure of the the strength
of the interaction.  Finally, we also measure the geometry of each
interaction through the values for $\mu$.

All the physical quantities described above can be used to describe
the three main stages in a bullet-like encounter. First, the sub-halo
crosses the virial radius of the host halo starting a head on
collision, $||\vec{r}||/R_{\rm vir}\approx 1$ and
$\mu\approx<0.0$. Second, as the sub-halo crosses for the first time
the center of the host halo $||\vec{r}||/R_{\rm vir} < 1.0$ and
$\mu>0.0$. Third, as the sub-halo reaches apogee and comes back to the
center of the halo $||\vec{r}||/R_{\rm vir} < 1.0$ and $\mu<0.0$. We use
this quantities in Section XXX to fully characterize the different
kind of interactions observed in the Bolshoi simulation.

>>>>

Since then, other examples of Bullet-like systems have been found; 
MACS J0025.4-1222 \citep{Bradac2008}, Abell 2744 \citep{Merten2011},
DLSCL J0916.2+2951 \citep{Dawson2012}, ZwCl 1234.0+02916
\citep{Dahle2013}. Recently \citep{Gastaldello} observed a baryonic-DM
displacement if $124\pm 20$ kpc in \bullg, a group-like system with a total
mass $2.4\pm 0.6 \times 10^{14}$\Msun. Systems of this mass are
$\sim10$ times more abundant than cluster systems in the mass range of
the Bullet Cluster $>10^{15}$\hMsun, this opens up the possibility of
observationally finding bullet groups in a fair amount to impose
constraints on $\Lambda$CDM.

@@

The Bullet-like configurations are composed by two dark matter
structures: the host halo and the dominant sub-halo. We describe the
kinematics of this configuration by the 
position and velocity vectors of the sub-halo in a frame of reference
where the main halo is at rest; thus
$\vec{v{}}=\vec{v}_{sub}-\vec{v}_{halo}$ and
$\vec{r}=\vec{r}_{sub}-\vec{r}_{halo}$, where the subscripts $host$
and $sub$ refer to the host and sub-halo in the frame of reference of
the simulation, respectively.

The angle between these two vectors can be characterized by, 
\begin{equation}
  \mu\equiv
  \cos(\theta)=\frac{\vec{v{}}\cdotp{}\vec{r}}{\left\|\vec{v}{}\right\|
    \left\|\vec{r}\right\|} .
 \end{equation} 
%
This encodes relevant information to describe the collision,
i.e. cases of $|\mu|\approx 1$ can be considered as head-on collisions
while $|\mu|\approx 0$ describes a grazing trajectory.

The geometrical configuration can be further described by the
following quantities. The circular velocity of each component, $V_{\rm
  c,host}$ for the host and $V_{\rm c, sub}$ for the sub-halo and the
size of the host halo $R_{\rm vir}$.

As a zero-th order approximation, in this paper we work with three
quantities that are available from observations of Bullet-like
systems. The projected distance between two dominant DM clumps, the
projected distance between the DM-baryonic clumps and the ratio of the galaxies
luminosities associated to the DM clumps.

From the simulation point of view, the first quantity can be
translated into the 2D projected values of $||\vec{r}||$ and its value
relative to the virial radius $D_{\rm   off}= ||\vec{r}||_{\rm
  2D}/R_{\rm vir}$. The last quantity can be approximated by ratio of
the circular velocities of the two clumps $V_{\rm c,sub}/V_{\rm
  c,host}$ which under conditions of constant mass-to-light ratios
should be equal to the square root of the luminosity ratio. The second
quantity is not directly available from a DM-only simulation but, as we
show in the Results section, can be  inferred from the available
information.

In order to gain better insight, we use two
quantities that are not readily available from observations but can be
measured in the simulation. The first one is the sub-structure
velocity as a fraction of the host's circular velocity, $||\vec{v}||/V_{\rm c, host}$, as a measure of the the strength of the
merger.  The second is $\mu$ to measure the geometry (head-on
vs. grazing) of each collision.

The physical quantities described above can be used to describe
the three main stages in a bullet-like encounter. First, the sub-halo
crosses the virial radius of the host halo starting a head on
collision, $||\vec{r}||/R_{\rm vir}\approx 1$ and
$\mu\approx<0.0$. Second, as the sub-halo crosses for the first time
the center of the host halo $||\vec{r}||/R_{\rm vir} < 1.0$ and
$\mu>0.0$. Third, as the sub-halo reaches apogee and comes back to the
center of the halo $||\vec{r}||/R_{\rm vir} < 1.0$ and $\mu<0.0$. We use
this quantities in the enxt section to fully characterize the different
kind of interactions observed in the Bolshoi Run,
